---
title: "SET Analyses - full pin time series; partial MH"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: false
        code_folding: 'hide'
        toc: yes
        toc_depth: 4
        toc_float: 
            collapsed: false
        fig_width: 7
        fig_height: 9
---


## Setup; data wrangling 

```{r setup_analyses, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


### Loading libraries and datasets

```{r importAndFormat_analyses}
# open up important libraries
library(tidyverse)
library(lubridate)
library(knitr)
library(broom)
library(agricolae)
library(gridExtra)
library(plotly)
library(EnvStats)
library(nerrSET)
# library(multcompView)

# load the datasets
load("SET_data_files.R")

```


The code above loaded a series of data frames created from the `Rates.Rmd` file. Contents are the following:  

+  Raw data:
    +  `markers_long`  
    +  `sets_long`  
+  Calculated rates:
    +  `allRates` - pin and MH rates, averaged by SET   
    +  `rates_pins` - raw pin rates; 36 per SET    
    +  `rates_MHs` - raw MH rates; 3 per SET  

    
### Adding site names to the datasets

Sites are as follows:  

-  __CLMA = Cladium:__
    - CLMAJ-1 
    - CLMAJ-2 
    - CLMAJ-3
-  __PANN = Panne:__
    - PANNE-1
    - PANNE-2
    - PANNE-3
-  __JULO = Lower Juncus:__
    - JURO-1
    - JURO-2
    - JURO-3
-  __JUUP = Upper Juncus:__
    - JURO-4
    - JURO-5
    - JURO-6
-  __SPAL = Spartina:__
    - SPAL-1
    - SPAL-2
    - SPAL-3
    
    


The code below uses the `nerrSET` package (on my machine only) to add site names to the data frames.  

```{r}
allRates <- attach_site_names(allRates, "SET")
markers_long <- attach_site_names(markers_long, "SET")
sets_long <- attach_site_names(sets_long, "SET")
rates_pins <- attach_site_names(rates_pins)
rates_MHs <- attach_site_names(rates_MHs)
```


Here, we check the heads (and one tail) of the data frames, to make sure site got added in correctly. 

```{r checkHeads_analyses}
# check the heads of the data frames
# kable(head(allRates, 3), caption="head of allRates (calculated data)", align='c')
# kable(tail(allRates, 3), caption="tail of allRates (calculated data)", align='c')
# kable(head(markers_long, 3), caption="head of markers_long (QCd raw data)", align='c')
# kable(head(sets_long, 3), caption="head of sets_long (QCd raw data)", align='c')
# kable(head(rates_pins, 3), caption="head of rates_pins (calculated data)", align='c')
# kable(head(rates_MHs, 3), caption="head of rates_MHs (calculated data)", align='c')
```



## Subsetting

We only want to look at the most recent laying of feldspar for each MHplot. This should help us look at subsidence.


Eventually, will calculate annualized rates of change for each. For now, I just want to get them subsetted and on the same graph.


```{r}
markers_long_lastlaying <- markers_long %>%
    group_by(SET, MHplot) %>%
    filter(ref_date >= max(ref_date))
```


## Change since 'time 0'


Don't need to calculate this for markers because that's what the reading means. Do need to average it per MHplot though.

For pins, this was already calculated earlier in a previous script, and is the field `totalchange` in this data frame.

So here, I'm just summarizing things.

```{r}
# get an average and SE for each SET/date
sets_long_summ <- sets_long %>%
    ungroup() %>%
    group_by(site, SET, date) %>%
    summarize(meanChange = mean(totalchange, na.rm=TRUE),
              seChange = sd(totalchange, na.rm=TRUE)/sqrt(length(totalchange)),
              nChange = length(totalchange))

```

Averaged readings for a single MH plot by date; into an intermediate data frame.

Then averaged readings of MH plots per SET per date, into the summary data frame that will be used for plotting.  

This way, if different MHplots had different numbers of readings on a given date, they're not unevenly weighted. Each MHplot is even by averaging within plots first.

```{r}
markers_long_intermed <- markers_long_lastlaying %>%
    ungroup() %>%
    group_by(site, SET, MHplot, date) %>%
    summarize(meanSed = mean(value, na.rm=TRUE))

markers_long_summ <- markers_long_intermed %>%
    ungroup() %>%
    group_by(site, SET, date) %>%
    summarize(meanSed = mean(meanSed, na.rm=TRUE))
```



## Graph  

### SETs alone first


```{r, fig.width=8, fig.height=5}
ggplot(sets_long_summ) +
    geom_point(aes(x=date, y=meanChange, col=site), size=2, alpha=0.7) +
    facet_wrap(~site) +
    theme_bw() +
    ggtitle("Cumulative Change", subtitle = "pin heights over time \nsince first reading") +
    xlab("Date") +
    ylab("change (mm)")


# 
# ggplot() +
#     geom_point(data=markers_long_summ, aes(x=date, y=meanSed), col="gray22") +
#     geom_point(data=sets_long_summ, aes(x=date, y=meanChange, col=site), size=2, alpha=0.5) +
#     facet_wrap(~site, ncol=1) +
#     theme_bw() +
#     ggtitle("Cumulative Change", subtitle = "pin heights over time") +
#     xlab("Date") +
#     ylab("change (mm)")
```



### MHs alone

#### Just points

```{r}

ggplot(markers_long_intermed) +
    geom_point(aes(x=date, y=meanSed, col=as.factor(MHplot), shape=as.factor(MHplot)), size=2, alpha=0.6) +
    facet_wrap(~SET, ncol=3) +
    theme_bw() +
    ggtitle("Mean Sediment Height by MHplot", subtitle = "MHplots over time") +
    xlab("Date") +
    ylab("change (mm)")
```



#### Points and loess smooth

```{r}
ggplot(markers_long_intermed) +
    geom_point(aes(x=date, y=meanSed, col=as.factor(MHplot), shape=as.factor(MHplot)), size=3, alpha=0.5) +
    geom_smooth(aes(x=date, y=meanSed), se=FALSE) +
    facet_wrap(~SET, ncol=3) +
    theme_bw() +
    ggtitle("Mean Sediment Height by MHplot", subtitle = "MHplots over time") +
    xlab("Date") +
    ylab("change (mm)")


```



### Combined Plots

What about just adding the pin readings on top of those loess smooths?  

***
***

#### pins - linear regression; MHs - loess (1 per SET)
One loess per SET's MHs; linear regression for pins:

```{r}

# ungroup the pin readings, just in case
sets_long_summ <- ungroup(sets_long_summ) %>%
    group_by(SET)

ggplot() +
    geom_point(data = markers_long_intermed, aes(x=date, y=meanSed, col=as.factor(MHplot), shape=as.factor(MHplot)), size=2, alpha=0.5) +
    geom_smooth(data = markers_long_intermed, aes(x=date, y=meanSed), se = FALSE) +
    geom_point(data = sets_long_summ, aes(x=date, y=meanChange), col="navyblue", size=2.5) +
    geom_smooth(data = sets_long_summ, aes(x=date, y=meanChange), method = "lm", col="navyblue", se = FALSE, size = 1.5) +
    facet_wrap(~SET, ncol=3) +
    theme_bw() +
    ggtitle("Change over time", subtitle = "pins are navy; MHplots are lighter colors \nlinear regression on pins; loess on MHplots") +
    xlab("Date") +
    ylab("change (mm)")

```


***
***


#### pins - linear regression; MHs - loess (1 per MHplot; 3 per SET)

Same graph, but with a loess for each individual MHplot:

```{r}

ggplot() +
    geom_point(data = markers_long_intermed, aes(x=date, y=meanSed, col=as.factor(MHplot), shape=as.factor(MHplot)), size=2, alpha=0.5) +
    geom_smooth(data = markers_long_intermed, aes(x=date, y=meanSed, col=as.factor(MHplot)), se=FALSE) +
    geom_point(data = sets_long_summ, aes(x=date, y=meanChange), col="navyblue", size=2.5) +
    geom_smooth(data = sets_long_summ, aes(x=date, y=meanChange), method = "lm", se = FALSE, col="navyblue", size=1.5) +
    facet_wrap(~SET, ncol=3) +
    theme_bw() +
    ggtitle("Change over time", subtitle = "pins are navy; MHplots are lighter colors \nlinear regression on pins; loess on MHplots") +
    xlab("Date") +
    ylab("change (mm)")

```


***
***


#### pins - loess; MHs - loess (1 per MHplot; 3 per SET)

Same graph, but with a loess for each individual MHplot, and loess for pins as well:

```{r}

ggplot() +
    geom_point(data = markers_long_intermed, aes(x=date, y=meanSed, col=as.factor(MHplot), shape=as.factor(MHplot)), size=2, alpha=0.5) +
    geom_smooth(data = markers_long_intermed, aes(x=date, y=meanSed, col=as.factor(MHplot)), se=FALSE) +
    geom_point(data = sets_long_summ, aes(x=date, y=meanChange), col="navyblue", size=2.5) +
    geom_smooth(data = sets_long_summ, aes(x=date, y=meanChange), method = "loess", se = FALSE, col="navyblue", size=1.5) +
    facet_wrap(~SET, ncol=3) +
    theme_bw() +
    ggtitle("Change over time", subtitle = "pins are navy; MHplots are lighter colors \nloess on pins and MHplots") +
    xlab("Date") +
    ylab("change (mm)")

```

***
***



#### pins - loess; MHs - loess (1 per SET)

```{r}

ggplot() +
    geom_point(data = markers_long_intermed, aes(x=date, y=meanSed, col=as.factor(MHplot), shape=as.factor(MHplot)), size=2, alpha=0.5) +
    geom_smooth(data = markers_long_intermed, aes(x=date, y=meanSed), se=FALSE) +
    geom_point(data = sets_long_summ, aes(x=date, y=meanChange), col="navyblue", size=2.5) +
    geom_smooth(data = sets_long_summ, aes(x=date, y=meanChange), method = "loess", se = FALSE, col="navyblue", size=1.5) +
    facet_wrap(~SET, ncol=3) +
    theme_bw() +
    ggtitle("Change over time", subtitle = "pins are navy; MHplots are lighter colors \nloess on pins and MHplots") +
    xlab("Date") +
    ylab("change (mm)")

```


***
***



#### pins - linear regression; MHs - linear regression (1 per SET)

```{r}

ggplot() +
    geom_point(data = markers_long_intermed, aes(x=date, y=meanSed, col=as.factor(MHplot), shape=as.factor(MHplot)), size=2, alpha=0.5) +
    geom_smooth(data = markers_long_intermed, aes(x=date, y=meanSed), method = "lm", se=FALSE) +
    geom_point(data = sets_long_summ, aes(x=date, y=meanChange), col="navyblue", size=2.5) +
    geom_smooth(data = sets_long_summ, aes(x=date, y=meanChange), method = "lm", se = FALSE, col="navyblue", size=1.5) +
    facet_wrap(~SET, ncol=3) +
    theme_bw() +
    ggtitle("Change over time", subtitle = "pins are navy; MHplots are lighter colors \nlinear regression on pins and MHplots") +
    xlab("Date") +
    ylab("change (mm)")

```


***
***



#### pins - linear regression; MHs - linear regression (1 per MHplot; 3 per SET)

```{r}

ggplot() +
    geom_point(data = markers_long_intermed, aes(x=date, y=meanSed, col=as.factor(MHplot), shape=as.factor(MHplot)), size=2, alpha=0.5) +
    geom_smooth(data = markers_long_intermed, aes(x=date, y=meanSed, col = as.factor(MHplot)), method = "lm", se=FALSE) +
    geom_point(data = sets_long_summ, aes(x=date, y=meanChange), col="navyblue", size=2.5) +
    geom_smooth(data = sets_long_summ, aes(x=date, y=meanChange), method = "lm", se = FALSE, col="navyblue", size=1.5) +
    facet_wrap(~SET, ncol=3) +
    theme_bw() +
    ggtitle("Change over time", subtitle = "pins are navy; MHplots are lighter colors \nlinear regression on pins and MHplots") +
    xlab("Date") +
    ylab("change (mm)")

```