---
title: "Marker_Horizons"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: true
        code_folding: 'hide'
        fig_width: 6
        fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Setup and Data Prep

Mostly this is turning numbers into the correct format (e.g. `Date`), generating site names, and setting up rows and columns the way I want them.

```{r, message=FALSE, warning=FALSE}
library(reshape2)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)

# setwd("C:/Users/kimberly.cressman/Desktop/Main Docs/Sentinel_Sites")

# format feldspar application data
feldspar <- read.csv("feldspar_apps.csv")
# glue site.platform and plat together
feldspar <- mutate(feldspar, site.platform.plat = paste0(site.platform, "_", plat))
# put application dates in date format
feldspar$feldspar_1 <- mdy(feldspar$feldspar_1)
feldspar$feldspar_2 <- mdy(feldspar$feldspar_2)


# read marker horizon data; glue site, platform, plat, and rep together into one identifier; gather into a long dataset
markers <- read.csv("marker_horizons.csv")
markers <- markers %>%
    mutate(site.platform.plat.rep = paste0(site.platform, "_", plat, "_", rep)) %>%
    gather(key = date, value = sed_height, -site.platform.plat.rep, -site.platform, -plat, -rep)

# remove the 'X' before the date and put it in date format
markers$date <- sub('.', '', markers$date) 
markers$date <- mdy(markers$date)

# give it sites:
markers <- mutate(markers, site = substr(site.platform, 1, 4))

# need to split upper and lower juncus
# JURO1-3 is lower
# 4-6 is upper
upper <- c("JURO_4", "JURO_5", "JURO_6")
lower <- c("JURO_1", "JURO_2", "JURO_3")
markers$site[markers$site.platform %in% upper] <- "UPJU"
markers$site[markers$site.platform %in% lower] <- "LOJU"


# make NAs something that will show up on the graph. -10?
markers$sed_height_noNA <- markers$sed_height
markers$sed_height_noNA[is.na(markers$sed_height_noNA)] <- -10
```


### See the data we're working with

First and last three rows of the data frame.

Variables:

* __site__ and __site.platform__ are the same as in the SETs file (e.g. CLMA and CLMAJ-1, respectively). 
* __plat__ is the number (out of 3) of the sediment core. 
* __rep__ is a letter identifier (A-D) for the four replicate measurements within that sediment core.
* __site.platform.plat.rep__ is all of that strung together as the identifier for a single measurement.
* __date__ is self-explanatory.
* __sed_height__ is the measurement of sediment in mm above the feldspar marker horizon.
* __sed_height_noNA__ replaces any missing measurements in the _sed_height_ column with -10 so they'll be visible on the graph.

```{r}
library(knitr)
kable(head(markers, 3))
kable(tail(markers, 3))
```


### Basic graph of sediment heights by date; NAs are included as -10s

```{r, warning=FALSE, message=FALSE, fig.height=9, fig.width=9}
ggplot(markers) +
    geom_point(aes(x=date, y=sed_height_noNA, col=as.factor(plat)), size=2, alpha=0.4) +
    facet_wrap(~site.platform, ncol=3) +
    theme_bw()



# make graphs by site.platform

# for(i in levels(markers$site.platform)) {
#     dat <- markers[markers$site.platform == i,]
#     feldspardates <- feldspar[feldspar$site.platform == i, "feldspar_2"]
#     g <- ggplot(dat) +
#         geom_point(aes(x=date, y=sed_height_noNA, col=as.factor(rep)), size=3, alpha=0.7) +
#         geom_vline(feldspardates)
#         facet_wrap(~site.platform + plat, ncol=1) +
#         theme_bw()
#     print(g)
# }

```


### Looking at plats (with error bars, +/- 1 standard error) within each site

Everything above got autoscaled and that made things kind of hard to see. Here it is by platform, showing each plat plus or minus one standard error. NAs were ignored in these graphs.

```{r, message=FALSE, warning=FALSE, fig.height=9, fig.width=7.5}
# group by site.platform.plat and summarize:
plats <- markers %>%
    mutate(site.platform.plat = paste0(site.platform, "_", plat)) %>%
    group_by(site.platform.plat, date) %>%
    dplyr::summarize(sed_mean = mean(sed_height, na.rm=TRUE), sed_stdev = sd(sed_height, na.rm=TRUE), sed_stderr = sed_stdev/sqrt(length(sed_height)))

# give it sites:
plats <- mutate(plats, 
                site = substr(site.platform.plat, 1, 4), 
                site.platform=substr(site.platform.plat, 1, nchar(site.platform.plat)-2), 
                plat=substr(site.platform.plat, nchar(site.platform.plat), nchar(site.platform.plat)))

# need to split upper and lower juncus
# JURO1-3 is lower
# 4-6 is upper
upper <- c("JURO_4", "JURO_5", "JURO_6")
lower <- c("JURO_1", "JURO_2", "JURO_3")
plats$site[plats$site.platform %in% upper] <- "UPJU"
plats$site[plats$site.platform %in% lower] <- "LOJU"
plats$site <- factor(plats$site)

# plot that

for(i in levels(plats$site)) {
    dat <- plats[plats$site == i,]

    g <- ggplot(dat) +
        geom_point(aes(x=date, y=sed_mean, col=as.factor(plat)), size=3, alpha=0.8) +
        geom_errorbar(aes(x = date, 
                          ymin=sed_mean - sed_stderr, 
                          ymax=sed_mean + sed_stderr, 
                          col=as.factor(plat))) +
        facet_wrap(~site.platform, ncol=1) +
        theme_bw()
    print(g)
}


```

