---
title: "MHs from the start"
author: "Kim Cressman"
date: "July 18, 2018"
mainfont: Arial
output: 
    html_document:
        css: C:/Users/kimberly.cressman/Documents/R/win-library/3.5/summarytools/includes/stylesheets/summarytools.css
        toc: true
        toc_depth: 4
        toc_float: yes
        code_folding: hide 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

__Initial RMD setup__
<br>

```{r setup_analyses, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

__Loading libraries and data sets__
<br>
```{r importAndFormat_analyses}
# open up important libraries
library(tidyverse)
library(lubridate)
library(knitr)
library(RColorBrewer)
library(nerrSET)
library(paletteer)
library(plotly)
library(skimr)
library(summarytools)
library(ggbeeswarm)
```

```{r}
# load the datasets
load("SET_data_files.R")

# get rid of everything but markers_long and sets_long
rm(list = c('allRates', 'rates_MHs', 'rates_pins'))
```


# Subset the data frame  

Get rid of any MH measurements on horizons laid after 3/2/2012.  


```{r}
markers2 <- markers_long %>%
    group_by(SET, MHplot) %>%
    filter(ref_date <= "2012-03-02")
```


## Attach Site names  

```{r}
markers2 <-  attach_site_names(markers2)
```


# Graph all the raw values  


Make a function to loop through the sites in the data frame  

```{r}
stns <- unique(markers2$site)

# create a container for our output
output <- htmltools::tagList()
        
        
for(i in 1:length(stns)) {
    # subset data
    dat <- markers2 %>%
        filter(site == stns[i])
    # generate title
    maintitle <- as.character(stns[i])
    # make a plot
    p <- ggplot(dat, aes(x = date, 
                         y = value, 
                         group = reading,
                         text = paste("Date:", as.character(date),
                                      "\nreading:", value,
                                      "\nfeldspar applied:", as.character(ref_date)))) +
        geom_point(size = 2, 
                   alpha = 0.5,
                   col = "darkslategray") +
        facet_grid(MHplot~SET, scales = "free_y") +
        ggtitle(maintitle, subtitle = "rows are MHplots; columns are SETs") +
        # scale_color_brewer("BuGn", direction = -1) +
        theme_bw() +
        theme(legend.position = "none")
    
    # output the plot
    output[[i]] <- as_widget(ggplotly(p, tooltip = c("text")))
}
```

```{r}
output
```


# Incremental Changes  

Calculate incremental change: first have to summarize by date, then subtract and calculate.

```{r}
markers3 <- markers2 %>%
    group_by(site, SET, MHplot, date) %>%
    summarize(plotmean = mean(value, na.rm = TRUE)) %>%
    mutate(inc_change = plotmean - lag(plotmean, 1),
           plotdatediff = as.numeric(date - lag(date, 1)),
           inc_rate = (inc_change / plotdatediff) * 365.25)
```


For informational purposes, let's summarize this data frame.

```{r, results='asis'}
markers3 %>% 
    ungroup() 

print(dfSummary(markers3, graph.magnif = 0.75), method = "render")
```

```{r}
markers3b <- markers3 %>%
    ungroup()

print(dfSummary(markers3b, graph.magnif = 0.75), method = "render")
```


By site  

```{r}
# by site
markers3c <-  markers3 %>% 
    ungroup() %>% 
    group_by(site) %>%
    
print(dfSummary(markers3c, graph.magnif = 0.75), method = "render")
```


## summarize incremental rates  

### by MHplot  

```{r}
markers4 <- markers3 %>%
    ungroup() %>%
    group_by(site, SET, MHplot) %>%
    summarize(mean_inc_change = mean(inc_change, na.rm = TRUE),
              sd_inc_change = sd(inc_change, na.rm = TRUE),
              mean_inc_rate = mean(inc_rate, na.rm = TRUE),
              sd_inc_rate = sd(inc_rate, na.rm = TRUE))

markers4b  <- markers4 %>%  
    ungroup() %>% 
    group_by(site) 

print(dfSummary(markers4b, graph.magnif = 0.75), method = "render")
    
knitr::kable(markers4b, align = "c")
```

### by SET  

```{r}
markers5 <- markers4 %>%
    ungroup() %>%
    group_by(site, SET) %>%
    summarize(mean_inc_change = mean(mean_inc_change, na.rm = TRUE),
              sd_inc_change = sd(sd_inc_change, na.rm = TRUE),
              mean_inc_rate = mean(mean_inc_rate, na.rm = TRUE),
              sd_inc_rate = sd(sd_inc_rate, na.rm = TRUE))

print(dfSummary(markers5, graph.magnif = 0.75), method = "render")

knitr::kable(markers5, align = "c")
```


### by site  

```{r}
markers6 <- markers5 %>%
    ungroup() %>%
    group_by(site) %>%
    summarize(mean_inc_change = mean(mean_inc_change, na.rm = TRUE),
              sd_inc_change = sd(sd_inc_change, na.rm = TRUE),
              mean_inc_rate = mean(mean_inc_rate, na.rm = TRUE),
              sd_inc_rate = sd(sd_inc_rate, na.rm = TRUE))

print(dfSummary(markers6, graph.magnif = 0.75), method = "render")

knitr::kable(markers6, align = "c")
```

# plot again  

```{r}
stns <- unique(markers3$site)

# create a container for our output
changeoutput <- htmltools::tagList()
rateoutput <- htmltools::tagList()
        
        
for(i in 1:length(stns)) {
    # subset data
    dat <- markers3 %>%
        filter(site == stns[i])
    # generate title
    maintitle <- as.character(stns[i])
    # make a plot
    p <- ggplot(dat, aes(x = date, 
                         y = inc_change, 
                         group = MHplot,
                         col = as.factor(MHplot),
                         text = paste("Date:", 
                                      as.character(date),
                                      "\nChange since last reading:", 
                                      round(inc_change, 2),
                                      "\nDays since last reading:", 
                                      plotdatediff,
                                      "\nRate since last reading:", 
                                      round(inc_rate, 2)))) +
        geom_point(size = 2, 
                   alpha = 0.5) +
        facet_wrap(~SET, scales = "free_y", ncol = 1) +
        ggtitle(maintitle, subtitle = "rows are MHplots; columns are SETs") +
        # scale_color_brewer("BuGn", direction = -1) +
        theme_bw() +
        theme(legend.position = "none")
    
    # output the plot
    changeoutput[[i]] <- as_widget(ggplotly(p, tooltip = c("text")))
    
    # make another plot, for rates
    # make a plot
    q <- ggplot(dat, aes(x = date, 
                         y = inc_rate, 
                         group = MHplot,
                         col = as.factor(MHplot),
                         text = paste("Date:", 
                                      as.character(date),
                                      "\nChange since last reading:", 
                                      round(inc_change, 2),
                                      "\nDays since last reading:", 
                                      plotdatediff,
                                      "\nRate since last reading:", 
                                      round(inc_rate, 2)))) +
        geom_point(size = 2, 
                   alpha = 0.5) +
        facet_wrap(~SET, scales = "free_y", ncol = 1) +
        ggtitle(maintitle, subtitle = "rows are MHplots; columns are SETs") +
        # scale_color_brewer("BuGn", direction = -1) +
        theme_bw() +
        theme(legend.position = "none")
    
    # output the plot
    rateoutput[[i]] <- as_widget(ggplotly(q, tooltip = c("text")))
}
```

## Plots of raw change since last reading  

```{r, fig.width = 7, fig.height = 9}
changeoutput
```


## Plots of rate since last reading  

```{r, fig.width = 7, fig.height = 9}
rateoutput
```


# Incremental changes by pin  

```{r}
sets <- attach_site_names(sets_long)

sets2 <- sets %>%
    ungroup() %>%
    group_by(site, SET, arm, pin) %>%
    mutate(inc_change = value - lag(value, 1),
           datediff = as.numeric(date - lag(date, 1)),
           inc_rate = (inc_change / datediff) * 365.25)

print(dfSummary(sets2, graph.magnif = 0.75), method = "render")
```


## average that per date for each SET  

```{r}
sets3 <- sets2 %>% 
    ungroup() %>% 
    group_by(site, SET, date) %>% 
    summarize(mean_inc_change = mean(inc_change, na.rm = TRUE),
              sd_inc_change = sd(inc_change, na.rm = TRUE),
              mean_inc_rate = mean(inc_rate, na.rm = TRUE),
              sd_inc_rate = sd(inc_rate, na.rm = TRUE))

print(dfSummary(sets3, graph.magnif = 0.75), method = "render")
```

### plot incremental change by date for sets  

```{r}
ggplot(sets3, aes(x = date, y = mean_inc_rate)) +
    geom_point() +
    facet_wrap(~SET, ncol = 3) +
    theme_bw()
```



# MH and pin incremental changes together  


make a data frame with  marker incremental change by SET by date   

```{r}
markers7 <- markers3 %>% 
    ungroup() %>% 
    group_by(site, SET, date) %>% 
    summarize(mean_inc_change = mean(inc_change, na.rm = TRUE),
              sd_inc_change = sd(inc_change, na.rm = TRUE),
              mean_inc_rate = mean(inc_rate, na.rm = TRUE),
              sd_inc_rate = sd(inc_rate, na.rm = TRUE))

print(dfSummary(markers7, graph.magnif = 0.75), method = "render")
```


sets3 has pin incremental change by SET by date  

put them on a graph together   

```{r, fig.width = 9, fig.height = 7}
p <- ggplot() +
    geom_point(data=sets3, aes(x = date, y = mean_inc_rate, col = "pins")) +
    geom_point(data=markers7, aes(x = date, y = mean_inc_rate, col = "markers")) +
    scale_color_manual(values = c("pins" = "red", "markers" = "blue")) +
    facet_wrap(~SET, ncol = 3) +
    theme_bw()

ggplotly(p)
```


Connect the dots:
```{r}
# p <- ggplot() +
#     geom_point(data=sets3, aes(x = date, y = mean_inc_rate, col = "pins")) +
#     geom_line(data=sets3, aes(x = date, y = mean_inc_rate, col = "pins")) +
#     geom_point(data=markers7, aes(x = date, y = mean_inc_rate, col = "markers")) +
#     geom_line(data=markers7, aes(x = date, y = mean_inc_rate, col = "markers"))
#     scale_color_manual(values = c("pins" = "red", "markers" = "blue")) +
#     facet_wrap(~SET, ncol = 3) +
#     theme_bw()
# 
# ggplotly(p)
```


```{r}
p <- ggplot() +
    geom_col(width = 2, data=sets3, aes(x = date, y = mean_inc_rate, col = "pins")) +
    geom_col(width = 2, data=markers7, aes(x = date, y = mean_inc_rate, col = "markers")) +
    scale_color_manual(values = c("pins" = "red", "markers" = "blue")) +
    facet_wrap(~SET, ncol = 3) +
    theme_bw()

ggplotly(p)
```


zoom in on the y-axis. some of these incremental rates are pretty crazy and I want to see the more moderate ones.  

```{r}
p <- ggplot() +
    geom_col(width = 2, data=sets3, 
             aes(x = date, y = mean_inc_rate, col = "pins"),
             position = "dodge") +
    geom_col(width = 2, data=markers7, 
             aes(x = date, y = mean_inc_rate, col = "markers"),
             position="dodge") +
    scale_color_manual(values = c("pins" = "red", "markers" = "blue")) +
    scale_y_continuous(limits = c(-30, 30)) +
    facet_wrap(~SET, ncol = 3) +
    theme_bw()

p
```

```{r, fig.width = 9, fig.height = 7}
p <- ggplot() +
    geom_point(data=sets3, aes(x = date, y = mean_inc_rate, col = "pins")) +
    geom_point(data=markers7, aes(x = date, y = mean_inc_rate, col = "markers")) +
    scale_color_manual(values = c("pins" = "red", "markers" = "blue")) +
    scale_y_continuous(limits = c(-30, 30)) +
    facet_wrap(~SET, ncol = 3) +
    theme_bw()

ggplotly(p)
```




```{r, fig.width = 9, fig.height = 7}

q <- ggplot() +
    geom_quasirandom(data=sets3, aes(x = "pins", y = mean_inc_rate, col = "pins")) +
    geom_quasirandom(data=markers7, aes(x = "markers", y = mean_inc_rate, col = "markers")) +
    scale_color_manual(values = c("pins" = "red", "markers" = "blue")) +
    facet_wrap(~SET, ncol = 3) +
    theme_bw()

ggplotly(q, tooltip = c('mean_inc_rate'))
```

# Summary of incremental change rates by SET  

## Marker Horizons (accretion)  

Make a summary table - average mean incremental rates of change for all dates at a SET

```{r}
markers8 <- markers7 %>% 
    ungroup() %>% 
    group_by(SET) %>% 
    summarize(grand_mean_inc_change = mean(mean_inc_change, na.rm = TRUE),
              grand_mean_inc_rate = mean(mean_inc_rate, na.rm = TRUE))
kable(markers8, align = "c", caption = "Mean rates by SET for Marker Horizons")
```


## SETs  

```{r}
sets4 <- sets3 %>% 
    ungroup() %>% 
    group_by(SET) %>% 
    summarize(grand_mean_inc_change = mean(mean_inc_change, na.rm = TRUE),
              grand_mean_inc_rate = mean(mean_inc_rate, na.rm = TRUE))
kable(sets4, align = "c", caption = "Mean rates by SET for elevation change")
```

