---
title: "SET/MH Rate Generation"
output:
  word_document:
    toc: yes
    toc_depth: '4'
  pdf_document:
    toc: yes
    toc_depth: 4
  html_document:
    code_folding: hide
    keep_md: no
    toc: yes
date: "`r Sys.Date()`"
---


## Setup; data import and formatting  

Reading in most updated csv files; they are in wide format and this code gathers them into long format.  

```{r setup_rates, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r importAndFormat_rates}
# open up important libraries
library(tidyverse)
library(lubridate)
library(knitr)
library(broom)
library(plotly)


### pins
# read in; wide format
sets_wide <- read.csv("sets_wide_UPDATED.csv") %>%
    mutate(date = mdy(date)) 
# gather, so it's one row per pin. group by SET, date, arm.
sets_long <- sets_wide %>%
    gather(key="pin", value="value", -SET, -date, -arm) %>%
    group_by(SET, date, arm)
# create a summary dataset from sets_long, which is already grouped by SET-date-arm
sets_summByArm <- sets_long %>%
    summarize(meanPinHt = mean(value, na.rm=TRUE),
              sdPinHt = sd(value, na.rm=TRUE),
              sePinHt = sdPinHt/(sqrt(length(value))))


### markers    
# read in; wide format
markers_wide <- read.csv("markers_UPDATED.csv") %>%
    mutate(date = mdy(date), ref_date = mdy(ref_date))
# gather; one row per reading
markers_long <- markers_wide %>%
    gather(key="reading", value="value", -SET, -MHplot, -date, -ref_date) %>%
    group_by(SET, MHplot, date) %>%
    arrange(SET, MHplot, date, reading)

```

****

## A peek at all the data tables that we're working with

***


### SETs

```{r pindfHeads}

kable(head(sets_wide, 3), caption = "Pin readings - wide format", align='c')
kable(head(sets_long, 3), caption = "Pin readings - long format", align='c')
# kable(head(sets_summByArm, 3), caption = "Pin readings - summary by arm position")

```

<a href="#top">Back to top</a> 

***
***

### Marker Horizons

```{r markerdfHeads}

kable(head(markers_wide, 3), caption = "MH readings - wide format", align='c')
kable(head(markers_long, 3), caption = "MH readings - long format", align='c')

```


<a href="#top">Back to top</a> 


****
****


## Rates galore



### SET readings

##### by individual pin

```{r setRatesByPin}

models <- sets_long %>%
    group_by(SET, arm, pin) %>%  
    do(mod = lm(value ~ date, data=.))

modelcoef <- tidy(models, mod, conf.int=TRUE, conf.level=0.95) # this gives the intercept, slope, and p values, among others; conf.int=TRUE, conf.level=0.95 gives the 95% confidence interval

# get rid of intercept row,
# calculate mm/yr for all the terms,
# round mm/yr to 3 digits,
# get rid of the mm/day stuff
rates_pins <- modelcoef %>%
    filter(term == "date") %>%
    mutate(rate_mm.yr = round(estimate*365.25, 3),
           se_mm.yr = round(std.error*365.25, 3),
           CIlow_mm.yr = round(conf.low*365.25, 3),
           CIhigh_mm.yr = round(conf.high*365.25, 3),
           p.value = round(p.value, 4),
           statistic = round(statistic, 3)) %>%
    select(SET, arm, pin, rate_mm.yr, se_mm.yr, CIlow_mm.yr, CIhigh_mm.yr, statistic, p.value)


modelsummary <- glance(models, mod) # this gives r^2 and adjusted r^2, and p values, among others

kable(head(rates_pins), align='c')
kable(head(modelsummary), digits=3, align='c')

# write.csv(rates_pins, "pin_rates.csv", row.names = FALSE)
# write.csv(modelsummary, "pin_regressions.csv", row.names = FALSE)

```

<a href="#top">Back to top</a> 


#### histograms of pin rate data distribution

```{r SEThistograms, fig.width=6, fig.height=15}

# calculate a data frame of mean per SET
# to generate vertical lines on the histograms
mean_pinrate <- rates_pins %>%
    group_by(SET) %>%
    summarize(mean_pins = mean(rate_mm.yr, na.rm=TRUE))

a <- ggplot(rates_pins) +
    geom_histogram(aes(rate_mm.yr), color='darkslategray', fill='cadetblue3') +
    geom_vline(aes(xintercept=mean_pins), mean_pinrate, color="red3", lwd=1) +
    facet_grid(SET ~.) +
    theme_bw() +
    ggtitle("Histograms of elevation change rates by SET", subtitle="red line represents mean rate for SET")

ggplotly(a)


```


<a href="#top">Back to top</a> 



##### summarized by arm position

_Note: Because I combine tables at the end, I'm not showing individual tables of rates here. I have checked them to verify the combined table is correct. If anyone wants to see the intermediate tables, the "# kable" line can be uncommented and run._ 

```{r setRatesByArm}

rates_pins_summByArm <- rates_pins %>%
    ungroup() %>%
    group_by(SET, arm) %>%
    summarize(pin_change_mean = round(mean(rate_mm.yr, na.rm=TRUE), 2),
              pin_change_sd = round(sd(rate_mm.yr, na.rm=TRUE), 3),
              pin_change_se = round(pin_change_sd/sqrt(length(rate_mm.yr)), 3))

rates_mean_perarm <- rates_pins_summByArm %>%
    select(SET, arm, pin_change_mean) %>%
    spread(key = arm, value = pin_change_mean)

# kable(rates_mean_perarm, caption="Mean rate of change (mm/yr) per arm position")
```


<a href="#top">Back to top</a> 


#### summarized by SET


First by averaging all 36 pins' rates together


```{r setRatesBySET}
# average of 36 pins

rates_pins_summBySET <- rates_pins %>%
    ungroup() %>%
    group_by(SET) %>%
    summarize(pin_change_mean = round(mean(rate_mm.yr, na.rm=TRUE), 2),
              pin_change_sd = round(sd(rate_mm.yr, na.rm=TRUE), 3),
              pin_change_se = round(pin_change_sd/sqrt(length(rate_mm.yr)), 3))


# kable(rates_pins_summBySET, caption="Summary statistics of rate of change (mm/yr) per SET, when averaging all pins together")

# write.csv(rates_pins_summBySET, "pin_rates_bySET.csv", row.names=FALSE)

```


Also calculated by averaging the rates of the 4 arms. As expected, mean was the same as above; sd and se were different. Table is not shown here but can be generated.

```{r setRatesByArm-SET}
# average of 4 arms

rates_SET_avgByArm <- rates_pins_summByArm %>%
    ungroup() %>%
    group_by(SET) %>%
    summarize(pin_change_avg = round(mean(pin_change_mean, na.rm=TRUE), 2),
              pin_change_sd = round(sd(pin_change_mean, na.rm=TRUE), 3),
              pin_change_se = round(pin_change_sd/sqrt(length(pin_change_mean)), 3))

# kable(rates_SET_avgByArm, caption = "Summary statistics of rate of change (mm/yr) per SET, when averaging arms together")

```


<a href="#top">Back to top</a> 

****

### Marker Horizons

```{r MHratesByCore}

# need a column for number of days between feldspar application and reading date
markers_long <- mutate(markers_long, daysSinceLaying = as.numeric(date - ref_date))

models <- markers_long %>%
    group_by(SET, MHplot) %>%  
    do(mod = lm(value ~ daysSinceLaying, data=.))

modelcoef <- tidy(models, mod, conf.int=TRUE, conf.level=0.95) # this gives the intercept, slope, and p values, among others; conf.int=TRUE, conf.level=0.95 gives the 95% confidence interval

# get rid of intercept row,
# calculate mm/yr for all the terms,
# round mm/yr to 3 digits,
# get rid of the mm/day stuff
rates_MHs <- modelcoef %>%
    filter(term == "daysSinceLaying") %>%
    mutate(rate_mm.yr = round(estimate*365.25, 3),
           se_mm.yr = round(std.error*365.25, 3),
           CIlow_mm.yr = round(conf.low*365.25, 3),
           CIhigh_mm.yr = round(conf.high*365.25, 3),
           p.value = round(p.value, 4),
           statistic = round(statistic, 3)) %>%
    select(SET, MHplot, rate_mm.yr, se_mm.yr, CIlow_mm.yr, CIhigh_mm.yr, statistic, p.value)


modelsummary <- glance(models, mod) # this gives r^2 and adjusted r^2, and p values, among others

kable(head(rates_MHs), align='c')
kable(head(modelsummary), digits=3, align='c')

# write.csv(rates_MHs, "MH_rates.csv", row.names = FALSE)
# write.csv(modelsummary, "pin_regressions.csv", row.names = FALSE)

```

<a href="#top">Back to top</a> 


#### histograms of accretion rate data distribution


```{r MHhistograms, fig.width=6, fig.height=15}

# calculate a data frame of mean per SET
# to generate vertical lines on the histograms
mean_mhrate <- rates_MHs %>%
    group_by(SET) %>%
    summarize(mean_mh = mean(rate_mm.yr, na.rm=TRUE))


a <- ggplot(rates_MHs) +
    geom_histogram(aes(rate_mm.yr), color='darkslategray', fill='tomato3') +
    geom_vline(aes(xintercept=mean_mh), mean_mhrate, color="navyblue", lwd=1) +
    facet_grid(SET ~.) +
    theme_bw() +
    ggtitle("Histograms of Marker Horizon change rates by SET", subtitle="blue line represents mean rate for SET")

ggplotly(a)

```


<a href="#top">Back to top</a> 


#### summarized by SET

_Again, not showing this table, because this information is in the final table._

```{r MHRatesBySET}
# average of 36 pins

rates_MHs_summBySET <- rates_MHs %>%
    ungroup() %>%
    group_by(SET) %>%
    summarize(MH_change_mean = round(mean(rate_mm.yr, na.rm=TRUE), 2),
              MH_change_sd = round(sd(rate_mm.yr, na.rm=TRUE), 3),
              MH_change_se = round(MH_change_sd/sqrt(length(rate_mm.yr)), 3))


# kable(rates_MHs_summBySET, caption="Summary statistics of rate of accretion (mm/yr) per SET, when averaging all 3 MHplots together")

# write.csv(rates_MHs_summBySET, "MH_rates_bySET.csv", row.names=FALSE)

```


****

## Combined pin and MH data

### Table of rates

```{r rateCombine}

allRates <- full_join(rates_pins_summBySET, rates_MHs_summBySET)

kable(allRates, align='c')

# write.csv(allRates, "rates_allBySET.csv", row.names=FALSE)

```


<a href="#top">Back to top</a> 


### bar graph of pin and MH rates by site

```{r rateGraph, fig.width=7, fig.height=5}

# need to make the data frame long

# allRates_long <- allRates %>%
#     gather(key = "param", value = "value", -SET) %>%
#     mutate(type = strsplit(param, "_"))

a <- ggplot(allRates) +
    geom_col(aes(x=SET, y=pin_change_mean, fill="pins"), col="darkslategray", alpha=0.5) +
    geom_col(aes(x=SET, y=MH_change_mean, fill="MH"), col="darkslategray", alpha=0.6) +
    geom_errorbar(aes(x=SET, ymin=pin_change_mean - 2*pin_change_se, ymax=pin_change_mean + 2*pin_change_se), col="navyblue") +
    geom_errorbar(aes(x=SET, ymin=MH_change_mean - 2*MH_change_se, ymax=MH_change_mean + 2*MH_change_se), col="red3") +
    theme_bw() +
    theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ggtitle("Rates of change by SET \n+/- 2 standard errors \n(approximately a 95% confidence interval)") +
  ylab("long-term change (mm)")
    
ggplotly(a)    

```


<a href="#top">Back to top</a> 



### faceted plot of pin changes  


Averaged per SET. First converted to "change since time 0" for each pin, then all changes averaged per SET per date.

```{r pinChangePlot, fig.width=7, fig.height=10}

# convert to change since time 0 for each pin
sets_long <- sets_long %>%
    group_by(SET, arm, pin) %>%
    arrange(SET, arm, pin, date) %>%
    mutate(totalchange = value - value[1])

# the 2 lines below load code for the 'stat_smooth_func', which annotates facets of a ggplot with the regression equation and r^2s. I opted to comment this out because the slope is in mm/day, which is hard to understand.
# library(devtools)
# source_gist("524eade46135f6348140", filename="ggplot_smooth_func.R")

a <- ggplot(ungroup(sets_long), aes(x=date, y=totalchange)) +
    geom_point(aes(col=arm), alpha=0.6) +
    geom_smooth(method="lm", se=FALSE) +
    # stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
    facet_wrap(~SET, ncol=3) +
    ggtitle("Changes in each pin since first measurement", subtitle = "colored by arm position, with one overall linear regression per SET") +
    labs(y="change since first reading (mm)", x="Measurement Date") +
    theme_bw()

ggplotly(a)

```


<a href="#top">Back to top</a> 


### faceted plot of MH changes  


Averaged per SET

```{r MHchangePlot, fig.width=8, fig.height=15}

markers_long <- markers_long %>%
    mutate(yearsSinceLaying = daysSinceLaying/365.25)


a <- ggplot(ungroup(markers_long)) +
    geom_point(aes(x=yearsSinceLaying, y=value, col=as.factor(ref_date)), size=2, alpha=0.5) +
    geom_smooth(aes(x=yearsSinceLaying, y=value), method="lm") +
    facet_grid(SET~MHplot) +
    ggtitle("Accretion rates by plot") +
    labs(x="Years since feldspar application", y="total change (mm)") +
    theme_bw()

ggplotly(a)

```


<a href="#top">Back to top</a> 


***
***

## Export the data frames, to be used in a later analysis script

This will create an R object named `SET_data_files.R`. Contents are the following data frames:  

+  Raw data:
    +  `markers_long`  
    +  `sets_long`  
+  Calculated rates:
    +  `allRates` - pin and MH rates, averaged by SET   
    +  `rates_pins` - raw pin rates; 36 per SET    
    +  `rates_MHs` - raw MH rates; 3 per SET  
    
    
_Note: This is commented out so it won't run every time; it can be run whenever there is a change that makes it necessary._
    
```{r dataExport}
# last updated 12/28 after adding columns above (years since feldspar application in markers_long and change since time 0 in sets_long)

# save(markers_long, sets_long, allRates, rates_pins, rates_MHs, file="SET_data_files.R")

```

