---
title: "SET Analyses - NAVD88"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: false
        code_folding: 'hide'
        toc: yes
        toc_float: true
        toc_depth: 4
---

__Initial RMD setup__
<br>

```{r setup_analyses, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

__Loading libraries and data sets__
<br>
```{r importAndFormat_analyses}
# open up important libraries
library(tidyverse)
library(lubridate)
library(knitr)
library(broom)
library(agricolae)
library(gridExtra)
library(plotly)
library(EnvStats)
# library(multcompView)

# load the datasets
load("SET_data_files.R")

```


# Add NAVD88 information into the SET data frame  

There is a spreadsheet called 'NAVD88.csv' that has NAVD88 elevations of each SET table in it. These were derived from _[duration]_ RTK occupations on _[some date]_. (2011?)

The data frame we want to work with here is `sets_long`. Here is the top of that data frame:
<br>
```{r}
kable(head(sets_long))
```

<br>

## Read in NAVD88 data  

Here we read in and look at the head of the NAVD88 sheet:
<br>
```{r}
navd88 <- read.csv("NAVD88.csv")
kable(head(navd88))
```

<br>

## Calculate NAVD88 offset for each SET, then NAVD88-adjusted surface elevations  

<br>
A correction factor can be calculated as: `SET_NAVD88 + Adapter_NAVD88 - 0.756`  

0.756m is the length of the pins used to measure surface elevation.  

Here I use the above formula to calculate an offset for each SET, and add it into the `navd88` data frame. I then add the calculated _offset_ value into the `sets_long` data frame, and subtract it from the pin value (which has been converted to meters), to calculate a navd88-adjusted surface elevation, named *elevation_navd88*.

<br>
```{r}
navd88 <- navd88 %>%
    mutate(offset = SET_NAVD88 + Adapter_NAVD88 - 0.756)
kable(head(navd88))

sets_long_navd88 <- left_join(sets_long, navd88, by = c("SET" = "Site")) %>%
    select(-SET_NAVD88, -Adapter_NAVD88) %>%
    mutate(value_m = value/1000,
           elevation_navd88 = value_m + offset)
kable(head(sets_long_navd88))
```

<br>

### Quick plots to see what's going on  

There has been no aggregation yet - this is all 36 pins per SET.  
<br>

```{r}
ggplot(sets_long_navd88) +
    geom_point(aes(x = date, y = elevation_navd88, color = SET)) +
    theme_minimal() +
    ggtitle("NAVD-88 adjusted pin heights")
```

<br>
Facet by site to zoom in a bit.
<br>
```{r}
ggplot(sets_long_navd88) +
    geom_point(aes(x = date, y = elevation_navd88, color = SET)) +
    facet_wrap(~SET, ncol = 3, scales = "free_y") +
    theme_minimal() +
    ggtitle("NAVD-88 adjusted pin heights")    
```

<br>
<br>

# Aggregate by SET  

Here I'm averaging all 36 pins within a SET to get a single, NAVD-88 adjusted, mean elevation for the SET on each date.
<br>

```{r}
sets_navd88_mean <- sets_long_navd88 %>%
    group_by(SET, date) %>%
    summarize(elevation_mean = mean(elevation_navd88, na.rm = TRUE))
kable(head(sets_navd88_mean))
```

< br>
Another plot to see how it looks now:
<br>
```{r}
ggplot(sets_navd88_mean) +
    geom_point(aes(x = date, y = elevation_mean, color = SET)) +
    theme_minimal() +
    ggtitle("NAVD88-adjusted mean surface elevation by SET")
```

<br>
<br>

## Add in column for site
<br>


# Sea Level Rise stuff

<br>


## Rate of SLR


The rate of sea level rise (SLR) at the Dauphin Island Tide gauge through 2016 was 3.5 mm/yr. This gets updated every year and in 2017 was 3.61. More details can be found here: 
https://tidesandcurrents.noaa.gov/sltrends/sltrends_station.shtml?stnid=8735180 


Because R works in days, we need to convert 3.5 mm/yr into a rate per day. And because the NAVD88 stuff is in meters, we need this to become meters per day. The math for that is:

3.5 * 365.25 / 1000

The 365.25 accounts for leap years.

<br>


## Sea Level values  



### Tidal Datums relative to MLLW  

Tidal datums were obtained from https://tidesandcurrents.noaa.gov/datums.html?units=1&epoch=0&id=8740166&name=Grand+Bay+NERR%2C+Mississippi+Sound&state=MS on 5/11/2018. 


Per page 7 of the benchmark sheets found here: https://tidesandcurrents.noaa.gov/benchmarks.html?id=8740166  , these datums were calculated based on readings at a tide station at the Bayou Heron Boat Ramp from March to August 2011.  

Tidal datums in meters, relative to MLLW:  

+  MHHW 0.485  
+  MHW 0.456  
+  MSL 0.236  
+  MLW 0.039  
+  MLLW 0.000  
+  NAVD88 0.183


### Tidal Datums relative to NAVD88  

These are obtained by subtracting 0.183 from all the other values, and these will be used in the graphs.  

+  MHHW 0.302  
+  MHW 0.273  
+  MSL 0.053  
+  MLW -0.144  
+  MLLW -0.183  
+  NAVD88 0.000 


Adding `geom_abline` to the graphs with x-axis as the year involved some calculations, per this stackoverflow answer: https://stackoverflow.com/questions/21946896/how-to-add-abline-in-ggplot2-with-x-axis-as-year

```{r}

# slope in m/day
slr_slope <- 3.5/365.25/1000

# and intercept, because time 0 is at 1970-01-01 and we want line to go through 0 on August 31, 2011, the end of the data period for datum calculation
slr_int <- 0.053 - slr_slope*as.integer(as.Date("2011-08-31"))
mhw_int <- 0.273 - slr_slope*as.integer(as.Date("2011-08-31"))
mlw_int <- -0.144 - slr_slope*as.integer(as.Date("2011-08-31"))
mhhw_int <- 0.302 - slr_slope*as.integer(as.Date("2011-08-31"))
mllw_int <- -0.183 - slr_slope*as.integer(as.Date("2011-08-31"))
```


# Graphs that include sea levels

## Lines but no shading


```{r}
ggplot(sets_navd88_mean, aes(x=date, y=elevation_mean)) +
    geom_point(aes(col=SET), alpha=0.5, size=3) +
    theme_minimal() +
    geom_abline(aes(slope = slr_slope, intercept = slr_int), col="blue") +
    geom_abline(aes(slope = slr_slope, intercept = mhw_int), col="blue") +
    geom_abline(aes(slope = slr_slope, intercept = mlw_int), col="blue") +
    geom_abline(aes(slope = slr_slope, intercept = mllw_int), col="blue", lty = 2) +
    geom_abline(aes(slope = slr_slope, intercept = mhhw_int), col="blue", lty = 2) +
    ggtitle("SETS through time relative to NAVD88") +
    ylab("elevation (m)") +
    ylim(-0.3,1.0)
```


__Facet because this is weird__

<br>

```{r}
ggplot(sets_navd88_mean, aes(x=date, y=elevation_mean)) +
    geom_point(aes(col=SET), alpha=0.5, size=3) +
    theme_minimal() +
    geom_abline(aes(slope = slr_slope, intercept = slr_int), col="blue") +
    geom_abline(aes(slope = slr_slope, intercept = mhw_int), col="blue") +
    geom_abline(aes(slope = slr_slope, intercept = mlw_int), col="blue") +
    geom_abline(aes(slope = slr_slope, intercept = mllw_int), col="blue", lty = 2) +
    geom_abline(aes(slope = slr_slope, intercept = mhhw_int), col="blue", lty = 2) +
    facet_wrap(~SET, ncol = 3) +
    ggtitle("SETS through time relative to NAVD88") +
    ylab("elevation (m)") +
    ylim(-0.3,1.0)

# free y-scales
ggplot(sets_navd88_mean, aes(x=date, y=elevation_mean)) +
    geom_point(aes(col=SET), alpha=0.5, size=3) +
    theme_minimal() +
    geom_abline(aes(slope = slr_slope, intercept = slr_int), col="blue") +
    geom_abline(aes(slope = slr_slope, intercept = mhw_int), col="blue") +
    geom_abline(aes(slope = slr_slope, intercept = mlw_int), col="blue") +
    geom_abline(aes(slope = slr_slope, intercept = mllw_int), col="blue", lty = 2) +
    geom_abline(aes(slope = slr_slope, intercept = mhhw_int), col="blue", lty = 2) +
    facet_wrap(~SET, ncol = 3, scales = "free_y") +
    ggtitle("SETS through time relative to NAVD88") +
    ylab("elevation (m)") 
```


__Sanity check: just horizonal lines; don't try to show SLR; just the tidal datums__


+  MHHW 0.302  
+  MHW 0.273  
+  MSL 0.053  
+  MLW -0.144  
+  MLLW -0.183  
+  NAVD88 0.000 

```{r}
ggplot(sets_navd88_mean, aes(x=date, y=elevation_mean)) +
    geom_point(aes(col=SET), alpha=0.5, size=3) +
    theme_minimal() +
    geom_hline(yintercept = 0.273, col="blue") +
    geom_hline(yintercept = 0.053, col="blue") +
    geom_hline(yintercept = -0.144, col="blue") +
    geom_hline(yintercept = 0.302, col="blue", lty = 2) +
    geom_hline(yintercept = -0.183, col="blue", lty = 2) +
    facet_wrap(~SET, ncol = 3) +
    ggtitle("SETS through time relative to NAVD88") +
    ylab("elevation (m)") +
    ylim(-0.3,1.0)
```


STILL NEED TO:  

+  Add in 'site' column for aggregating  
+  make the shaded graph

