---
title: "Interactive SET/MH graphs"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: true
        code_folding: 'hide'
        fig_width: 5
        fig_height: 3
---

***
***


These .csv files have been run through a few steps before getting to this point.  

The raw SET-pin data and Marker Horizon data spreadsheets were imported into R and changed into a long format. For SET-pin data, this happened in `SET_processing_and_summarizing.R`. That file was further run through `SETs.Rmd` to format correctly, etc.; then through `wide_format_conversion.R` to get it into wide format (individual SET tables and sample dates as row identifiers; arm-pin as column identifiers).

Marker horizons were read in through the `SETs plus MHs.Rmd` file, then run through `wide_format_conversion.R` to get into wide format with SET table and sample dates as row identifiers, and plat-rep (e.g. 1-A) as column names.

Once the files were in wide format, with only one row per site and date, I went through the field notebooks and corrected any date discrepancies.

This file will read the files with corrected dates back into R make interactive graphs using the `dygraphs` and/or `ggiraph` packages.

Note: `markers_with_references.csv` is the file with all correct dates and 0s for every re-laying of feldspar.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
library(lubridate)
library(broom)
library(knitr)
library(agricolae)
library(dygraphs)
library(xts)
library(htmltools)
library(ggplot2)
library(ggiraph)
```

### Read in data

```{r, message=FALSE, warning=FALSE}
sets <- read.csv("sets_wide_corrected.csv", stringsAsFactors = FALSE) %>%
    mutate(date = mdy(date)) %>%
    gather(key = arm.pin, value = pin_height, -site.platform, -date) %>%
    mutate(arm.pin = sub('.', '', arm.pin)) %>%
    arrange(site.platform, date)

    
markers <- read.csv("markers_with_references.csv", stringsAsFactors = FALSE) %>%
    mutate(date = mdy(date), plat=paste0("plat", plat)) %>%
    group_by(site.platform, plat, date) %>%
    mutate(sed_height = mean(c(A, B, C, D), na.rm=TRUE)) %>%
    select(site.platform, plat, date, sed_height) %>%
    distinct() %>%
    spread(key = plat, value = sed_height)

feldspar_relays <- read.csv("markers_with_references.csv", stringsAsFactors = FALSE) %>%
    mutate(date = mdy(date), 
           ref_date = mdy(ref_date),
           plat=paste0("relay", plat), 
           ref = 0) %>%
    select(site.platform, plat, ref_date, ref) %>%
    distinct() %>%
    spread(key = plat, value = ref)
    
```


### Work up some summaries

#### Pin readings:  

*  group by site.platform and date; calculate mean, st deviation, st error  
*  subtract first reading at site from all subsequent readings. NOTE that `pin_change$change` is elevation change, in mm, relative to the first reading, which is on or around 2-29-2012 at all sites.  
*  add column for overall site names (e.g. UPJU = Upper Juncus = JURO-4, JURO-5, and JURO-6)

```{r}
# group by platform and date, then calculate mean and standard deviation
pin_summary <- dplyr::summarize(group_by(sets, site.platform, date), mean = mean(pin_height, na.rm=TRUE), stddev = sd(pin_height, na.rm=TRUE), stderr = stddev/sqrt(length(pin_height))) 

# subtract first reading at a platform from all subsequent readings
# split, apply, combine
pin_split <- split(pin_summary, pin_summary$site.platform) %>%
    lapply(., FUN = function(x) {mutate(x, "change" = mean - mean[1])})
pin_change <- do.call(rbind, pin_split) %>%
    mutate(., site = substr(site.platform, 1, 4))

# need to split upper and lower juncus
# JURO1-3 is lower
# 4-6 is upper
upper <- c("JURO-4", "JURO-5", "JURO-6")
lower <- c("JURO-1", "JURO-2", "JURO-3")
pin_change$site[pin_change$site.platform %in% upper] <- "UPJU"
pin_change$site[pin_change$site.platform %in% lower] <- "LOJU"

# clean up the temporary files
rm(pin_summary)
rm(pin_split)
rm(lower)
rm(upper)
```

NAVD-88 adjustments:

```{r}
# set up a NAVD88 data frame
NAVD88 <- read.csv("NAVD88.csv")
# calculate the adjustment factor that will need to be applied to pin heights:
### total height (set + adapter + reading) - pin length (0.756) = NAVD88 location of marsh surface
NAVD88 <- mutate(NAVD88, totalfactor = SET_NAVD88 + Adapter_NAVD88 - 0.756)

# calculate NAVD-adjusted pin heights
# first add the adjustment factor to the data frame
pin_change$navdadj <- NAVD88$totalfactor[match(pin_change$site.platform, NAVD88$Site)]

# then calculate the adjusted value
# BEWARE - the units are different. turn this all into meters.
pin_change <- mutate(pin_change, meanadj = (mean/1000) + navdadj)
```



#### Marker Horizons:

```{r}
# give it sites:
markers <- mutate(markers, site = substr(site.platform, 1, 4))

# need to split upper and lower juncus
# JURO1-3 is lower
# 4-6 is upper
upper <- c("JURO_4", "JURO_5", "JURO_6")
lower <- c("JURO_1", "JURO_2", "JURO_3")
markers$site[markers$site.platform %in% upper] <- "UPJU"
markers$site[markers$site.platform %in% lower] <- "LOJU"

rm(upper)
rm(lower)

```

Marker Horizon Sample Dates

```{r}
# kable(markerdates)
```

SET Sample Dates

```{r}
# kable(setdates)
```


### Glue together marker horizon and pin change data

First have to generate mean accretion, stdev, and sterr for each platform

```{r, message=FALSE, warning=FALSE}
# make the punctuation the same in both data frames (turn marker summary ones from CLMAJ_1 to CLMAJ-1 etc.); specify that it's a factor
markers$site.platform <- factor(sub("_", "-", markers$site.platform))

feldspar_relays$site.platform <- factor(sub("_", "-", feldspar_relays$site.platform))

# join
dat_almost <- full_join(pin_change, markers, by=c("site.platform", 'date'))

dat_all <- full_join(dat_almost, feldspar_relays, by=c("site.platform", "date" = "ref_date"))
```

Update some names; select only pin and MH readings; turn it into a time series object

```{r}
dat_small <- dat_all %>%
    mutate(pins = change, markers1 = plat1, markers2 = plat2, markers3 = plat3) %>%
    select(site.platform, date, pins, markers1, markers2, markers3, relay1, relay2, relay3)
dat_small <- xts(dat_small, order.by = dat_small$date)

```

***
***

### Dygraphs!

Commented out because I can fine-tune the plots from `ggiraph` more easily. (don't want to lose this code though!)

Needs a bit of cleanup, but notice you can hover over a point and see the values; you can also change the range to zoom in on an area. Because of how it's graphed, you can see when a series isn't continuous - which would indicate either NAs for MHs or mismatched dates between pin and MH measurements.


```{r, fig.width=7, fig.height=4}
### NOTE: I've changed the marker horizons since writing this. There is no longer an "average" per date; I'm using all three plats, so the code will need updating.


# # for loops and html widgets don't play nice together, so I've had to split the data frame, write my own function, and use lapply and htmltools to generate a plot for each site.platform


# # here's the function
# set_interact <- function(dat_split){
#     dygraph(data = dat_split, main = as.character(unique(dat_split$site.platform)), xlab="Date", ylab="mm") %>%
#         dySeries("pins", label = "SET pins", drawPoints = TRUE, pointSize = 4, color = "blue") %>%
#         dySeries("markers", label = "Marker Horizons", drawPoints = TRUE, pointSize = 4, color = "red") %>%
#         dyRangeSelector()
# }
# 
# # here's the data frame splitting
# sets_split <- split(dat_small, dat_small$site.platform)
# 
# # aaaaand lapply
# # which needs to be wrapped in htmltools, per Yihui
# # https://stackoverflow.com/questions/30509866/for-loop-over-dygraph-does-not-work-in-r
# toprint <- lapply(sets_split, set_interact)
# 
# # although we also need space between the graphs, so here's a solution
# # https://stackoverflow.com/questions/38109020/how-do-i-add-spacing-between-dygraphs-generated-in-apply-loop-in-rmarkdown
# 
# space <- list(htmltools::tags$h1("--------"), 
#               htmltools::tags$h2("--------"))
# 
# out <- c(rbind(space, toprint))
# 
# 
# # and print it out
# htmltools::tagList(out)


```


## Interactive plots using ggplot2 and ggiraph

+  Hovering over a point will show the date and the reading.  
+  You can also pan and zoom in on a graph by using the buttons that appear in its upper right corner when you hover over the plot area.
+  The blue points and lines are averages of SET pin data. I wanted this to be the most noticeable grouping because these graphs have a lot going on - so the points are solid, but mostly transparent, so you can see what's happening with the triangles.
+  The triangles are averages from each of 3 individual MH plots for each SET.
+  Filled-in triangles (at y=0) represent dates when feldspar was applied or reapplied. You will notice that on some dates, it was only reapplied at one of the three plots; this is what prevents me from averaging across all plots on all dates.  
+  __Breaks in the lines__ represent either missing data (most likely NAs on the MHs) OR __dates when SET and MH readings were done on different days__. Usually they are only separated by 1-3 days, but there is at least one instance when the two readings for a quarter were done more than a month apart.  



```{r, message=FALSE, warning=FALSE}


# make columns of date and measurement for the tooltip
# first, remake dat_small; not as a time-series object
dat_small <- dat_all %>%
    mutate(pins = change, markers1 = plat1, markers2 = plat2, markers3 = plat3) %>%
    select(site.platform, date, pins, markers1, markers2, markers3, relay1, relay2, relay3)

dat_small <- dat_small %>%
    mutate(pin_caption = paste0(date, ": ", round(pins,2), " mm"),
           mh1_caption = paste0(date, ": ", round(markers1, 2), " mm"),
           mh2_caption = paste0(date, ": ", round(markers2, 2), " mm"),
           mh3_caption = paste0(date, ": ", round(markers3, 2), " mm"))

# split the data
sets_split <- split(dat_small, dat_small$site.platform)


set_gginteract <- function(dat_split){
g <- ggplot(dat_split, aes(x=date)) +
    geom_point_interactive(aes(y=pins, tooltip=pin_caption, col="SET pins"), alpha = 0.3, size = 4, shape = 16) +
    geom_point_interactive(aes(y=markers1, tooltip=mh1_caption, col="Marker Horizons 1"), size = 3, shape = 2) +
    geom_point_interactive(aes(y=markers2, tooltip=mh2_caption, col="Marker Horizons 2"), size = 3, shape = 2) +
    geom_point_interactive(aes(y=markers3, tooltip=mh3_caption, col="Marker Horizons 3"), size = 3, shape = 2) +
    geom_point(aes(y=relay1, col="Marker Horizons 1"), size = 3, shape = 17, alpha = 0.8) +
    geom_point(aes(y=relay2, col="Marker Horizons 2"), size = 3, shape = 17, alpha = 0.8) +
    geom_point(aes(y=relay3, col="Marker Horizons 3"), size = 3, shape = 17, alpha = 0.8) +
    geom_line(aes(y=pins, col="SET pins")) +
    geom_line(aes(y=markers1, col="Marker Horizons 1"), alpha = 0.5) +
    geom_line(aes(y=markers2, col="Marker Horizons 2"), alpha = 0.5) +
    geom_line(aes(y=markers3, col="Marker Horizons 3"), alpha = 0.5) +
    theme_bw() +
    labs(title = as.character(unique(dat_split$site.platform)), x ="Date", y = "mm") +
    scale_color_manual(values = c("SET pins" = "blue", "Marker Horizons 1" = "red", "Marker Horizons 2" = "orange", "Marker Horizons 3" = "darkslategray"))

ggiraph(code = print(g + theme(legend.position='bottom')), width_svg = 10, tooltip_opacity = 0.8, zoom_max = 10)
}

# aaaaand lapply
# which needs to be wrapped in htmltools, per Yihui
# https://stackoverflow.com/questions/30509866/for-loop-over-dygraph-does-not-work-in-r
toprint <- lapply(sets_split, set_gginteract)

# don't really need the contortions for spacing that we had above, so just print it out
htmltools::tagList(toprint)
```

***
***
***

## Pull out only dates where SETs and MHs were both measured

I went back to the step where I joined the data frames (line 170 as I'm writing this) to only join dates that are in BOTH data frames.   

To do this: I used `dplyr::left_join` with `sets` as the left data frame and `markers` as the right. This pulled in MH data that matched SET measurement dates and excluded any MH data from other dates. On dates when SETs were measured but MHs were not, NAs were placed in the columns where matching MH data would have been. I eventually filtered out those rows, where SETs had been measured and MHs had not, by removing rows that had NAs in MH columns.  

NAs can be meaningful, and I did not want to remove rows that had legitimate NAs because of MHs not being found on a measurement date. So before I joined the data frames together, I turned "NA" values into "-9999". Then when I removed rows with NAs in certain columns, the __real__ NAs remained. After joining and filtering, I turned -9999s back into NAs.


```{r}

### transform the NAs:
markers2 <- markers 
markers2$plat1[is.na(markers2$plat1) == "TRUE"] <- -9999
markers2$plat2[is.na(markers2$plat2) == "TRUE"] <- -9999
markers2$plat3[is.na(markers2$plat3) == "TRUE"] <- -9999
    
# join and filter out the NAs in plat1 (because this will also get rid of NAs in plats 2 and 3):
dat_almost2 <- left_join(pin_change, markers2, by=c("site.platform", 'date')) %>%
    filter(!is.na(plat1))

# pull in feldspar relay dates
dat_samedates <- left_join(dat_almost2, feldspar_relays, by=c("site.platform", "date" = "ref_date"))

# turn -9999s back into NAs
dat_samedates$plat1[dat_samedates$plat1 == -9999] <- NA
dat_samedates$plat2[dat_samedates$plat2 == -9999] <- NA
dat_samedates$plat3[dat_samedates$plat3 == -9999] <- NA
```

### Make the interactive graphs  

Because these were based on SET data, that is present in every row, and the lines for pin readings are fully connected. __However__, that doesn't mean every SET reading is included here - particularly in 2013, there are long stretches without a point. Those are dates when SETs and MHs were done on different days.


Missing triangles on a date are true NAs - where MHs were meant to be measured on a date when SETs were also measured, but feldspar was not found.


Doing both things on the same date seems to have become the habit starting at the end of 2013.


Note - there is one possible QC problem I found from these graphs; I believe, based on these graphs and the pdf named `pin_heights_by_site`, that the measurements for SPAL-1 and SPAL-3 were switched on 8-5-2014. I've sent the information to our SET team and will update data as necessary.

```{r, message=FALSE, warning=FALSE}


# make columns of date and measurement for the tooltip
# first, remake dat_small; not as a time-series object
dat_small <- dat_samedates %>%
    mutate(pins = change, markers1 = plat1, markers2 = plat2, markers3 = plat3) %>%
    select(site.platform, date, pins, markers1, markers2, markers3, relay1, relay2, relay3)

dat_small <- dat_small %>%
    mutate(pin_caption = paste0(date, ": ", round(pins,2), " mm"),
           mh1_caption = paste0(date, ": ", round(markers1, 2), " mm"),
           mh2_caption = paste0(date, ": ", round(markers2, 2), " mm"),
           mh3_caption = paste0(date, ": ", round(markers3, 2), " mm"))

# split the data
sets_split <- split(dat_small, dat_small$site.platform)


set_gginteract <- function(dat_split){
g <- ggplot(dat_split, aes(x=date)) +
    geom_point_interactive(aes(y=pins, tooltip=pin_caption, col="SET pins"), alpha = 0.3, size = 4, shape = 16) +
    geom_point_interactive(aes(y=markers1, tooltip=mh1_caption, col="Marker Horizons 1"), size = 3, shape = 2) +
    geom_point_interactive(aes(y=markers2, tooltip=mh2_caption, col="Marker Horizons 2"), size = 3, shape = 2) +
    geom_point_interactive(aes(y=markers3, tooltip=mh3_caption, col="Marker Horizons 3"), size = 3, shape = 2) +
    geom_point(aes(y=relay1, col="Marker Horizons 1"), size = 3, shape = 17, alpha = 0.8) +
    geom_point(aes(y=relay2, col="Marker Horizons 2"), size = 3, shape = 17, alpha = 0.8) +
    geom_point(aes(y=relay3, col="Marker Horizons 3"), size = 3, shape = 17, alpha = 0.8) +
    geom_line(aes(y=pins, col="SET pins")) +
    geom_line(aes(y=markers1, col="Marker Horizons 1"), alpha = 0.5) +
    geom_line(aes(y=markers2, col="Marker Horizons 2"), alpha = 0.5) +
    geom_line(aes(y=markers3, col="Marker Horizons 3"), alpha = 0.5) +
    theme_bw() +
    labs(title = as.character(unique(dat_split$site.platform)), x ="Date", y = "mm") +
    scale_color_manual(values = c("SET pins" = "blue", "Marker Horizons 1" = "red", "Marker Horizons 2" = "orange", "Marker Horizons 3" = "darkslategray"))

ggiraph(code = print(g + theme(legend.position='bottom')), width_svg = 10, tooltip_opacity = 0.8, zoom_max = 10)
}

# aaaaand lapply
# which needs to be wrapped in htmltools, per Yihui
# https://stackoverflow.com/questions/30509866/for-loop-over-dygraph-does-not-work-in-r
toprint <- lapply(sets_split, set_gginteract)

# don't really need the contortions for spacing that we had above, so just print it out
htmltools::tagList(toprint)
```