---
title: "Parent File for running SET analyses"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: false
        code_folding: 'hide'
        toc: yes
---

# Background  

## This document  

Code is hidden but can be expanded  

Glimpses of tables are included so the reader can keep track of how data is formatted throughout the process



## The Sentinel Site Project  

Something here about overall project goals


### Sampling Design

Long-term monitoring, etc. 5 sites along a coastal transition transect, each with 3 Surface Elevation Tables (SETs). 36 pins per SET, measured quarterly from 2012-2016, then 2x/yr. 3 Marker Horizon plots per SET, measured quarterly from 2012-2016, then 2x/yr.


### Data processing  

Master spreadsheet --> long or wide formats  

Read in through _xxxx_ script, then processed.  

Note corrections that have been made to original data.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1: Rate calculations
SET pins for elevation change  

Marker Horizons for accretion

```{r child = "Rates.Rmd"}
```

***
***
***
***

# 2: Statistical Analyses
Inter-site comparisons

```{r child = "Analyses.Rmd"}
```

***
***
***
***

# Housekeeping  
Package info, etc.

```{r}
sessionInfo()
```

