---
title: "SET Analyses - NAVD88"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
    fig_width: 5.5
    fig_height: 6
  html_document:
    code_folding: hide
    keep_md: no
    toc: yes
    toc_depth: 4
    toc_float: yes
date: "`r Sys.Date()`"
---

__Initial RMD setup__
<br>

```{r setup_analyses, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

__Loading libraries and data sets__
<br>
```{r importAndFormat_analyses}
# open up important libraries
library(tidyverse)
library(lubridate)
library(knitr)
# library(broom)
# library(agricolae)
# library(gridExtra)
# library(plotly)
# library(EnvStats)
library(RColorBrewer)
# library(multcompView)
# source("functions_sentinel_sites.R") # replaced by my personal package 'nerrSET'
library(nerrSET)
library(rcartocolor) # for Antique color palette
library(pals)
library(paletteer)

# load the datasets
load("SET_data_files.R")

```

__Note to self about figure sizes__: When fiddling with graphics in the console, execute `windows()` to open up a new graphics device; then run the code for the plot; resize with the mouse; then find the dimensions by typing `dev.size("in")`.

<br>


# Add NAVD88 information into the SET data frame  

There is a spreadsheet called 'NAVD88.csv' that has NAVD88 elevations of each SET table in it. These were derived from _[duration]_ RTK occupations on _[some date]_. (2011?)

The data frame we want to work with here is `sets_long`. Here is the top of that data frame:
<br>
```{r}
kable(head(sets_long))
```

<br>

## Read in NAVD88 data  

Here we read in and look at the head of the NAVD88 sheet:
<br>
```{r}
navd88 <- read.csv("NAVD88.csv")
kable(head(navd88))
```

<br>

## Calculate NAVD88 offset for each SET, then NAVD88-adjusted surface elevations  

<br>
A correction factor can be calculated as: `SET_NAVD88 + Adapter_NAVD88 - 0.756`  

0.756m is the length of the pins used to measure surface elevation.  

Here I use the above formula to calculate an offset for each SET, and add it into the `navd88` data frame. I then add the calculated _offset_ value into the `sets_long` data frame, and subtract it from the pin value (which has been converted to meters), to calculate a navd88-adjusted surface elevation, named *elevation_navd88*.

<br>
```{r}
navd88 <- navd88 %>%
    mutate(offset = SET_NAVD88 + Adapter_NAVD88 - 0.756)
kable(head(navd88))

sets_long_navd88 <- left_join(sets_long, navd88, by = c("SET" = "Site")) %>%
    select(-SET_NAVD88, -Adapter_NAVD88) %>%
    mutate(value_m = value/1000,
           elevation_navd88 = value_m + offset)
kable(head(sets_long_navd88))
```

<br>

### Quick plots to see what's going on  

There has been no aggregation yet - this is all 36 pins per SET.  
<br>

```{r, fig.width = 6.35, height = 8.35}
# ggplot(sets_long_navd88) +
#     geom_point(aes(x = date, y = elevation_navd88, color = SET)) +
#     theme_minimal() +
#     ggtitle("NAVD-88 adjusted pin heights")
```

<br>
Facet by site to zoom in a bit.
<br>
```{r, fig.width = 7.1, fig.height = 8.2}
# ggplot(sets_long_navd88) +
#     geom_point(aes(x = date, y = elevation_navd88, color = SET)) +
#     facet_wrap(~SET, ncol = 3, scales = "free_y") +
#     theme_minimal() +
#     ggtitle("NAVD-88 adjusted pin heights")    
```

<br>
<br>

# Aggregate by SET  

Here I'm averaging all 36 pins within a SET to get a single, NAVD-88 adjusted, mean elevation for the SET on each date.
<br>

```{r}
sets_navd88_mean <- sets_long_navd88 %>%
    group_by(SET, date) %>%
    summarize(elevation_mean = mean(elevation_navd88, na.rm = TRUE))
kable(head(sets_navd88_mean))
```

< br>
Another plot to see how it looks now:
<br>
```{r, fig.width = 4.667, fig.height = 8}
# ggplot(sets_navd88_mean) +
#     geom_point(aes(x = date, y = elevation_mean, color = SET)) +
#     theme_minimal() +
#     ggtitle("NAVD88-adjusted mean surface elevation by SET")
```

<br>
<br>

## Add in column for site
<br>

This is where my `nerrSET` package comes in.

```{r}
sets_navd88_mean <- attach_site_names(sets_navd88_mean)
sets_navd88_mean <- mutate(sets_navd88_mean, site = fct_relevel(site, "CLMA", "PANN", "JUUP", "JULO", "SPAL"))
               
```

# Graphing with different color palettes  

```{r}
p <- ggplot(sets_navd88_mean) +
    geom_point(aes(x = date, 
                   y = elevation_mean, 
                   color = site,
                   shape = site), 
               size = 2, 
               alpha = 0.7) +
    # theme_minimal() +
    ggtitle("NAVD88-adjusted mean surface elevation by SET") + 
    # scale_y_continuous(limits = c(0.05, 0.45)) + 
    geom_smooth(aes(x = date, 
                    y = elevation_mean, 
                    color = site), 
                method = "lm", 
                # lty = 2,
                alpha = 0.6,
                lwd = 1,
                se = FALSE) +
    scale_shape_manual(values = c(16, 17, 16, 6, 15)) + 
    theme(axis.line = element_line(colour = "gray20", 
                                   size = 0.5, linetype = "solid"), 
          panel.grid.major = element_line(linetype = "blank"), 
          panel.grid.minor = element_line(linetype = "blank"), 
          panel.background = element_rect(fill = NA), 
          legend.key = element_rect(fill = NA), 
          legend.background = element_rect(fill = NA))

```

\newpage

## ggplot2 default  

```{r}
p
```

\newpage

## RCartoColor  

### Antique, but from paletteer package  

```{r}
p + scale_color_paletteer_d(rcartocolor, Antique)
```


\newpage

### Prism, from paletteer package  

```{r}
p + scale_color_paletteer_d(rcartocolor, Prism)
```

\newpage

## RColorBrewer  

### YlGnBu  

```{r}
p + scale_color_manual(values = brewer.pal(9, "YlGnBu")[4:8])
```
\newpage

### PuBuGn  

```{r}
p + scale_color_manual(values = brewer.pal(9, "PuBuGn")[4:8])
```
\newpage

### PuBu  

```{r}
p + scale_color_manual(values = brewer.pal(9, "PuBu")[4:8])
```

\newpage
### GnBu  

```{r}
p + scale_color_manual(values = brewer.pal(9, "GnBu")[4:8])
```

\newpage
### Greens  

```{r}
p + scale_color_manual(values = brewer.pal(9, "Greens")[4:8])
```


\newpage
### Greys  

```{r}
p + scale_color_manual(values = brewer.pal(9, "Greys")[4:8])
```

\newpage
### OrRd  

```{r}
p + scale_color_manual(values = brewer.pal(9, "OrRd")[4:8])
```

\newpage
### Reds  

```{r}
p + scale_color_manual(values = brewer.pal(9, "Reds")[4:8])
```

\newpage
### YlOrBr  

```{r}
p + scale_color_manual(values = brewer.pal(9, "YlOrBr")[4:8])
```

\newpage

## pals  

### tol  

```{r}
p + scale_color_paletteer_d(pals, tol)
```
\newpage

## cartography  

### multi.pal  

```{r}
p + scale_color_paletteer_d(cartography, multi.pal, dynamic = TRUE)
```
\newpage

### blue.pal  

```{r}
p + scale_color_paletteer_d(cartography, blue.pal, dynamic = TRUE)
```

\newpage
### green.pal  

```{r}
p + scale_color_paletteer_d(cartography, green.pal, dynamic = TRUE)
```
\newpage
### harmo.pal  

```{r}
p + scale_color_paletteer_d(cartography, harmo.pal, dynamic = TRUE)
```