---
title: "SET Regressions"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: true
        code_folding: 'hide'
        fig_width: 8
        fig_height: 6
---

__Note__: R code in this document is hidden by default, but the chunks used for each step that follows can be viewed by clicking "code" on the right side of the file.   
   
### Setup & Data Prep   
   
1. Loading data that was processed via `SET_processing_and_summarizing.R` script; available here: https://github.com/swmpkim/Sentinel_Sites/blob/master/SET_processing_and_summarizing.R   

2. Processing some NAVD88 information to calculate an adjustment factor for pin heights: NAVD88 location of marsh surface = total measured height (leveled SET + leveled adapter + pin reading) - pin length (0.756m)
    - __Note__: 0.756 as pin length needs to be verified with Will U.

3. Subtracting first pin reading (actually the average of 36 readings) at a platform (e.g., JURO-1) from all subsequent readings to generate change over time.

4. Creating a column for site:   
    -  __CLMA = Cladium:__
        - CLMAJ-1 
        - CLMAJ-2 
        - CLMAJ-3
    -  __PANN = Panne:__
        - PANNE-1
        - PANNE-2
        - PANNE-3
    -  __LOJU = Lower Juncus:__
        - JURO-1
        - JURO-2
        - JURO-3
    -  __UPJU = Upper Juncus:__
        - JURO-4
        - JURO-5
        - JURO-6
    -  __SPAL = Spartina:__
        - SPAL-1
        - SPAL-2
        - SPAL-3

5. Calculating NAVD-88 adjusted pin heights

```{r, message=FALSE, warning=FALSE}
library(reshape2)
library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)
# load group of files, generated from SET_summarizing.R
load("sets.group.R")

# set up a NAVD88 data frame
NAVD88 <- read.csv("NAVD88.csv")
# calculate the adjustment factor that will need to be applied to pin heights:
### total height (set + adapter + reading) - pin length (0.756) = NAVD88 location of marsh surface
NAVD88 <- mutate(NAVD88, totalfactor = SET_NAVD88 + Adapter_NAVD88 - 0.756)

# subtract first reading at a platform from all subsequent readings
# split, apply, combine
pin_split <- split(pin_summary, pin_summary$site.platform) %>%
    lapply(., FUN = function(x) {mutate(x, "change" = mean - mean[1])})
pin_change <- do.call(rbind, pin_split) %>%
    mutate(., site = substr(site.platform, 1, 4))

# need to split upper and lower juncus
# JURO1-3 is lower
# 4-6 is upper
upper <- c("JURO-4", "JURO-5", "JURO-6")
lower <- c("JURO-1", "JURO-2", "JURO-3")
pin_change$site[pin_change$site.platform %in% upper] <- "UPJU"
pin_change$site[pin_change$site.platform %in% lower] <- "LOJU"

# calculate NAVD-adjusted pin heights
# first add the adjustment factor to the data frame
pin_change$navdadj <- NAVD88$totalfactor[match(pin_change$site.platform, NAVD88$Site)]

# then calculate the adjusted value
# BEWARE - the units are different. turn this all into meters.
pin_change <- mutate(pin_change, meanadj = (mean/1000) + navdadj)
```
   
   
In the data frame `pin_change`, we now have the following columns:   
   
- __site.platform__ - the individual SET platform, e.g. CLMAJ-1
- __date__ - date of pin reading
- __mean__ - mean of 36 pin readings on that date
- __stddev__ - standard deviation of 36 pin readings on that date
- __stderr__ - standard error of the mean for 36 in readings on that date
- __site__ - site that the SET platform belongs to, e.g. CLMA
- __sdymin__ - ymin for plotting error bars using standard deviation (mean - stddev)
- __sdymax__ - ymax for plotting error bars using standard deviation (mean + stddev)
- __seymin__ - ymin for plotting error bars using standard error (mean - stderr)
- __seymax__ - ymax for plotting error bars using standard error (mean + stderr)
- __change__ - change (mm) since first reading
- __navdadj__ - adjustment factor to relate mean pin readings to NAVD88; calculated for each individual site.platform
- __meanadj__ - mean marsh elevation (m) relative to NAVD88 on the sample date: mean of 36 pin readings, converted to m, plus the adjustment factor for the individual site.platform
   
   
Here are the first and last 3 rows of the data frame we're working with:   
   
```{r}
library(knitr)
kable(head(pin_change, 3))
kable(tail(pin_change, 3))
```


***

### Regressions


__Change over time at each individual platform, plus regression lines, compared to SLR__

Faceted by site because it's a little too crowded otherwise. The line is the regression for the site (e.g. CLMA; n=3 platforms per site). Reminder: all sites start at 0.


Local sea level rise at Dauphin Island is 3.5 mm/yr (source:  https://tidesandcurrents.noaa.gov/sltrends/sltrends_station.shtml?stnid=8735180 ). That is the gray dashed line in the following graphs. The intercept was set so the line would go through y=0 on the same date where pin change is 0.


Here are the slopes (and other regression outputs) for each site. Slopes were calculated in mm/day by default; look to the right-most columns for mm/yr, which I calculated.

#### Important note

I screwed up parentheses in the calculations for the original tables I sent, so those mm/yr slopes and confidence intervals are INCORRECT. The following calculations have been corrected.

```{r}
models <- pin_change %>%
    group_by(site) %>%  
    do(mod = lm(change ~ date, data=.))

modelcoefbysite <- tidy(models, mod, conf.int=TRUE, conf.level=0.95) # this gives the intercept, slope, and p values, among others; conf.int=TRUE, conf.level=0.95 gives the 95% confidence interval

# make a slope per year for date estimate; NAs for intercept
modelcoefbysite$mm.yr <- ifelse(modelcoefbysite$term == "date", round((modelcoefbysite$estimate * 4 * 365 + 1)/4, 3), "NA")

# confidence intervals in mm/yr:
modelcoefbysite$conf.low.yr <- ifelse(modelcoefbysite$term == "date", round(modelcoefbysite$conf.low * 365.25, 3), "NA")
modelcoefbysite$conf.high.yr <- ifelse(modelcoefbysite$term == "date", round(modelcoefbysite$conf.high * 365.25, 3), "NA")

modelsummarybysite <- glance(models, mod) # this gives r^2 and adjusted r^2, and p values, among others

kable(modelcoefbysite)
kable(modelsummarybysite)
```


#### Let's see those regressions on a plot.


```{r}
# to get that into rise per day, accounting for leap year:
slr_slope <- 3.5/365.25
# and intercept, because time 0 is at 1970-01-01 and we want line to go through 0
slr_int <- 0 - slr_slope*as.integer(pin_change[1,]$date)

ggplot(pin_change, aes(x=date, y=change)) +
    geom_point(aes(col=site.platform), alpha=0.5, size=3) +
    geom_smooth(method="lm", aes(group=site, col=site), show.legend=FALSE) +
    geom_abline(slope=slr_slope, intercept = slr_int, lty=2, lwd=1.5, alpha=0.7) +
    theme_minimal() +
    facet_wrap(~site, nrow=2) +
    ggtitle("Cumulative Surface Elevation Change") +
    ylab("change (mm)")
```


#### Now go site by site and look at diagnostics, particularly residual plots.


Split the data frame so each site has its own.

```{r}
spal <- filter(pin_change, site=="SPAL")
pann <- filter(pin_change, site=="PANN")
upju <- filter(pin_change, site=="UPJU")
loju <- filter(pin_change, site=="LOJU")
clma <- filter(pin_change, site=="CLMA")

```


#### SPAL

Here's something interesting: running regressions for `change` and `mean` (mean pin height) against `date` give the same slope, as they should, but different standard errors, t values, and p-values. Also, different R^2s.

```{r}
fit <- lm(change ~ date, data=spal)
fit2 <- lm(mean ~ date, data=spal)
fit3 <- lm(I(meanadj * 1000) ~ date, data=spal)

summary(fit)
# summary(fit2)
summary(fit3)

plot(spal$date, residuals(fit), xlab="date", main="residuals from fit of change")
abline(0,0)

plot(spal$date, residuals(fit3), xlab="date", main="residuals from fit of NAVD- adjusted mean")
abline(0,0)

# plot(spal$date, residuals(fit2), xlab="date", main="residuals from fit of mean")
# abline(0,0)

```


#### Upper Juncus

```{r}
fit <- lm(change ~ date, data=upju)
fit2 <- lm(mean ~ date, data=upju)

summary(fit)
summary(fit2)

plot(upju$date, residuals(fit), xlab="date", main="residuals from fit of change")
abline(0,0)

plot(upju$date, residuals(fit2), xlab="date", main="residuals from fit of mean")
abline(0,0)

```

